Estimating the likelihood of student replication success after an initial failure to replicate

### 2024-03-06 manuscript/manuscript.Rmd
work on methods

@@

Students were required to think of reasons the original replication might not have worked, and address them if they could. A list of possible reasons and solutions LINK was given to students. Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

do we have something to say about the range of interventions that went into replications? e.g., larger sample size, manipulation check, etc.

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool). We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level. (See table TODO for post-exclusion sample sizes of original, replication, and rescues). TODO  should say something about average power or average sample size as a multiplier on original or something - basically, a summary stat beyond a pointer to the table.
Across the `r num_proj-1` Prolific studies, we spent $`r pluck(cost,1,1) |> round()`, for an average of $`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP. Disagreements were resolved through discussion. As a compliment to the overall subjective rating of success, we followed @boyce2023 in also doing a statistical comparison on one key measure of interest for each study.

 We also recorded the same set of potential correlates that were used in @boyce2023 for original, replication, and rescue (these were already rated for original and replication). These potential correlates included features of the original study including the subfield of the study (cognitive v social v other psychology), it's publication year,  experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, and how many items each participant saw, and whether there were open materials and open data.

@@

Not all rescues of small replications succeeded, however. @payne2008 was a study of the effects of sleep versus wake on memory consolidation that showed participants a number of images and then hours later (after either sleep or no sleep) measured their recall for parts of the images. The first replication struggled to recruit participants and only got 23 (the original had 48). The rescue attempted to recruit a larger sample (target 88), but due to difficulties getting participants to complete the second part of the experiment 12 hours after the first, the rescue only managed to recruit 23 people. The lesson here may be that sleep research is difficult to conduct online. However, an online replication by @denis2022 with a much larger sample found qualitatively similar results to @payne2008, albeit with a much smaller effect size. TODO confirm once denis is reanalysed

>>>>

```{r}
rel_size <- parsed_d |> filter(type %in% c("original", "rescue")) |> select(target_lastauthor_year, type, N)  |> 
  pivot_wider(names_from=type, values_from=N) |> mutate(multiplier=rescue/original)

lowest <- quantile(rel_size$multiplier, 0)
low <- quantile(rel_size$multiplier, .25)
mid <- quantile(rel_size$multiplier, .5)
m <- mean(rel_size$multiplier)
high <- quantile(rel_size$multiplier, .75)
highest <- quantile(rel_size$multiplier, 1)

@@

Students were required to think of reasons the original replication might not have worked, and address them if they could. A list of possible reasons and solutions TODO LINK was given to students. In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies. For instance, TODO rescue switched from using the CRT which is overused in online samples for the newer TODO. TODO rescue discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli. TODO @tarampi original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with press keyboard keys to indicate. TODO others to note?  Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool). We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level. The rescues had on average `r round(m,2)` times the original sample post-exclusion (median: `r round(mid,2)`, IQR: `r round(low,2)` - `r round(high,2)`, minimum: `r round(lowest, 2)`, maximum: `r round(highest,2)`, see TABLE TODO for all sample sizes). 
Across the `r num_proj-1` Prolific studies, we spent $`r pluck(cost,1,1) |> round()`, for an average of $`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP. Disagreements were resolved through discussion. As a compliment to the overall subjective rating of success, we followed @boyce2023 in also doing a statistical comparison on one key measure of interest for each study. In order to statistically compare the key measures, we needed effect sizes reported in the same way within each original-replication cluster. When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, replication, and rescue (these were already rated for original and replication). These potential correlates included features of the original study including the subfield of the study (cognitive v social v other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, and how many items each participant saw, and whether there were open materials and open data.

@@

Not all rescues of small replications succeeded, however. @payne2008 was a study of the effects of sleep versus wake on memory consolidation that showed participants a number of images and then hours later (after either sleep or no sleep) measured their recall for parts of the images. The first replication struggled to recruit participants and only got 23 (the original had 48). The rescue attempted to recruit a larger sample (target 88), but due to difficulties getting participants to complete the second part of the experiment 12 hours after the first, the rescue only managed to recruit 23 people. The lesson here may be that sleep research is difficult to conduct online. However, an online replication by @denis2022 reports qualitatively similar (but quantitatively smaller) results to @payne2008 on related but not identical measures.
